Interpretable Neural Networks using EAGGA

### 2025-03-13 paper/main.tex
add bullets Pareto optimality

\subsection{Multi-Objective Optimization}
- in most practical use cases one doesn't just want to optimise for performance but e.g. also for interpretability using some proxy metrics
- formal definition
\subsubsection{Pareto-optimality}
\subsubsection{A-priori}
- requires specifying trade-off between objectives a-priori
- simplest approach: different versions of \textbf{scalarization}
  * weighted sum of objective functions
    - $\min_{x\in\mathcal{X}} w_i f_i(x)$ with $\sum_{i=1}^k w_i=1$ and $w_i>0,\forall i=1,...,k$
    - \citep[p. 11]{10.1145/3610536}
  * Chebyshev: provably converges to Pareto front
\subsubsection{A-posteriori}
- problem a priori
  * either restrict search space to ``enforce our will'' (e.g. only use 50\% of features), optimise only prediction performance + use this
  * or leave search space unrestricted but adjust loss function to incorporate multiple objectives + take optimum from there
  * in either case: no knowledge of interplay between HP config + performance on all objectives
  * a-posteriori evaluates configs just as a-priori, but keeps track of multiple ``best'' (non-dominated) solutions -> makes relationship between HP config + objectives visible
- NSGA-II (MOEA)
  * non-dominated sorting
  * crowding distance
  -> give intuition why non-dom-sort (asc) + crowding dist (desc) makes sense for selecting individuals

>>>>

\subsection{Multi-Objective Optimization}
- in most practical use cases one doesn't just want to optimise for performance but e.g. also for interpretability using some proxy metrics
- formal definition
  * vector of objectives $c_1,c_2,...,c_m=\boldsymbol{c}:\Lambda\rightarrow\mathbb{R}^m$
  * goal: minimise vector
  * \citep[p. 11]{10.1145/3610536}
\subsubsection{Pareto-optimality}
- problem: usually conflicting objectives, i.e. minimisation of $\boldsymbol{c}$ along all dimensions not possible
- thus aim to find trade-off solutions of non-dominated points
- we say a point dominates another
  * if there is no other point strictly better in at least one dimension and better or equal in the remaining ones
  * formally: $\boldsymbol\lambda$ dominates $\boldsymbol\lambda'$ ($\boldsymbol\lambda\prec\boldsymbol\lambda'$) if and only if
    $\forall i\in\{1,...,m\}:c_i(\boldsymbol\lambda) \le c_i(\boldsymbol\lambda') \wedge \exists j\in\{1,...,m\}:c_j(\boldsymbol\lambda) < c_j(\boldsymbol\lambda')$
  * \citep[pp. 7f]{10.1145/3610536} + \citep[pp. 198f]{genetic_algos}
- Pareto set: set of nondominated points $\mathcal{P}:=\{\boldsymbol\lambda\in\Lambda|\nexists\boldsymbol\lambda'\in\Lambda\text{ s.t. }\boldsymbol\lambda'\prec\boldsymbol\lambda\}$
- Pareto front: image of nondominated points
- goal: find set of nondominated points $\hat{\mathcal{P}}$ approximating true Pareto set $\mathcal{P}$ well
\subsubsection{A-priori}
- requires specifying trade-off between objectives a-priori
- simplest approach: different versions of \textbf{scalarization}
  * weighted sum of objective functions
    - $\min_{x\in\mathcal{X}} w_i f_i(x)$ with $\sum_{i=1}^k w_i=1$ and $w_i>0,\forall i=1,...,k$
    - \citep[p. 11]{10.1145/3610536}
  * Chebyshev: provably converges to Pareto front
\subsubsection{A-posteriori}
- problem a priori
  * either restrict search space to ``enforce our will'' (e.g. only use 50\% of features), optimise only prediction performance + use this
  * or leave search space unrestricted but adjust loss function to incorporate multiple objectives + take optimum from there
  * in either case: no knowledge of interplay between HP config + performance on all objectives
  * a-posteriori evaluates configs just as a-priori, but keeps track of multiple ``best'' (non-dominated) solutions -> makes relationship between HP config + objectives visible
- NSGA-II (MOEA)
  * non-dominated sorting
  * crowding distance
  -> give intuition why non-dom-sort (asc) + crowding dist (desc) makes sense for selecting individuals
